\documentclass[english,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
\newenvironment{ltable}
  {\begin{landscape}\begin{center}\begin{threeparttable}}
  {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

\usepackage{ifthen} % Only add declarations when endfloat package is loaded
\ifthenelse{\equal{man}{\string jou}}{%
  \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
  \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
  \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
}{}%


% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}


\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Using Financial Report Release Dates In Straddle Strategies},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\setcounter{secnumdepth}{0}
\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}



\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Using Financial Report Release Dates In Straddle Strategies}

  \shorttitle{FINANCIAL REPORTS AND STRADDLES}


  \author{Nelson Corrocher, Adela Devarajan, \& Nikoloz Lelashvili}

  \def\affdep{{"", "", ""}}%
  \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{} Harrisburg University  }

  \authornote{
    \newcounter{author}
    Nelson Corrocher, M.B.A.\\
    Adela Devarajan\\
    Nikoloz Lelashvili

                      Correspondence concerning this article should be addressed to Nelson Corrocher. E-mail: \href{mailto:ncorrocher@my.harrisburgu.edu}{\nolinkurl{ncorrocher@my.harrisburgu.edu}}
                                    }


  \abstract{In the context of financial investment, creating a
``straddle''(Investopedia, 2003) close to the release date of the
financial statement from public traded companies has been used by many
fund managers for some time. The goal of this research is to measure the
if a systematic approach of such strategy using blue chip stock options
can indeed yield above-the-average returns.}
  \keywords{straddle options financial reports \\

    \indent Word count: 58
  }




\begin{document}

\maketitle



During a trading simulation in a graduate investment class at Boston
University in 2014, the winning strategy used a straddle close to the
release date of the 10-K of a certain public company. Such performance
raised the question of using a systematic investment in blue chip
straddles close to the companyies' financial statement release as a
potential strategy in which the returns would be above average. It is
currently known that many fund managers use such strategy as part of
their strategies' portfolio. However, since they use additional criteria
to decide which company stocks should they straddle. So far, no found
literature studied the expected returns of taking a systematic
investment approach of creating straddles on the most negotiated public
companies, the so called Blue Chips.\\
 The approach of measuring the returns invoves collecting stock prices
from Blue Chip companies and comparing their price variation close to
the release data of Quarterly and Yearly financial to their average
price variation. If the hypothesis that the two sets (variation close to
the release date and at a random date) is rejected, a simulation will be
ran on past data to see if a systematic approach could yield returns
greater than the passive index average, which could provide a
complementary strategy to the passive index investment.\\
 The data is going to collected from public available sources as it
could be easily replicated by others. The methodology is still being
defined based on the availability and quality of data.

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

No direct participants were involved in this research as the main inputs
for the analysis were financial market data public available.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-straddle}{}
Investopedia. (2003). Straddle definition. Retrieved from
\url{http://www.investopedia.com/terms/s/straddle.asp}





\end{document}
