---
title: "Assignment 5 - Part 1 - In Class Assignment"
author: "Nelson Corrocher"
date: "July 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# In Class Assigment

## Example 1

We use father.son dataset to investigate regressions and correlations, in a way to facilitate reproducibility

The code below is used to prepare the environment for this example. It uses Father.Son data, which is part of the "UsingR" package:
```{r env, include = TRUE}
if (!require("UsingR")) {
  install.packages("UsingR", repos="http://cran.rstudio.com/") 
  library("UsingR")
}
```
  
Next, we standardize and center both father and son heights, preparing the variables for analysis:
```{r std, tab.cap=""}
y <- (father.son$sheight - mean(father.son$sheight)) / sd(father.son$sheight)
x <- (father.son$fheight - mean(father.son$fheight)) / sd(father.son$fheight)

```
*Obs: We could have used the function scale instead*    

Finally, we check the correlation between the two variables:
```{r cor}
rho <- cor(x,y)
rho
```

Now we plot the correlation between these two variables graphically:
```{r plot}
library(ggplot2)
g <- ggplot(data.frame(x = x, y = y), aes(x = x, y = y)) ## Creates 'g', a ggplot variable

# The following two lines set up the point aesthetics of the variable
g = g + geom_point(size = 6, colour = "black", alpha = 0.2) 
g = g + geom_point(size = 4, colour = "salmon", alpha = 0.2)

# and them set up the limits of visible graph (if outliers are present, the visuals could be messed up without it)
g = g + xlim(-4,4) + ylim(-4,4)

# The next three lines create graph lines, one vertical, one horizontal and one with the slope = 1, all passing through the origin
g = g + geom_abline(intercept = 0, slope = 1)
g = g + geom_vline(xintercept = 0)
g = g + geom_hline(yintercept = 0)

# These create some a line that using the slope of the correlation and a second one with the inverse of the correlation
g = g + geom_abline(intercept = 0, slope = rho, size = 2)
g = g + geom_abline(intercept = 0, slope = 1/rho, size = 2)
g
```

## Example 2

Below we create variables representing Federal Reserve rates.  
*Obs: Data is available at http://www.federalreserve.gov/release/g19/20050805/* 
```{r usfed-preset}
year <- c(2000,2001,2002,2003,2004)
rate <- c(9.34, 8.50, 7.62, 6.93, 6.60)
```

Now we show the plot representing the previous variables:
```{r fedplot}
plot(year, rate, main = "Commercial Banks Interest Rates", sub = "http://www.federalreserve.gov/release/g19/20050805/")
```

And now we show the correlation between the two variables:
```{r fedcor}
cor(year, rate)
```
As we can see, the correlatio is close to -1, which in practice, in uncommon.


Now we create the model for rate(dependent variable) to year(independent variable)
```{r fedfit}
fit <- lm(rate ~ year) # The first Variable is the dependent and the second one is the independent
fit
```a

Understanding that composes the fit variable:
```{r defattr}
attributes(fit)
```

To check how good the model is, we now are going to check the residuals:
```{r fedres}
res <- rate - (fit$coefficients[[2]]*year+fit$coefficients[[1]])
res
```

and we plot the residuals:
```{r redresplo}
plot(year,res)
```

Here we plot the federal rates with the fit line
```{r plotfin}
plot(year,rate)
abline(fit)
```

And finally we get a summary on the model
```{r summ}
summary(fit)
```
